---
title: "Task1"
author: "Your name"
date: "10/01/2024"
format:
  html:
    embed-resources: true
editor: source
---

## Question 1 Atomic Types

Create the following variables related to your chosen object:

4 numeric values (e.g., population, temperature, size, rating)
3 integer values (e.g., number of moons, buildings, inhabitants)
2 logical values (e.g., has_water, is_inhabited)
3 character values (e.g., name, region, category)

My object is Poltava city

```{r}
# code for solving task

#numeric
temperature <- 14.3
meters_above_sea_level <- 140.0
auto <- 285.000
highest_building <- 44.0


# integer

population <- 279593
area <- 103 #m2
shoppin_mall <- 3

#logical

has_water <-TRUE
has_autobus_station <- TRUE

#character

name <- 'Poltava'
region <- 'Poltavsky'
category <- 'city'

```


## Question 2 Vectors

Question description

Create three different vectors:

Numeric vector combining all your numeric and integer variables.
(Use c().)

Sequence vector from any number between 10–50 to 100, Choose step between 2 and 10.
(Use seq() or :.)

Logical vector of length 10 with random TRUE/FALSE values using sample().

```{r}
# code for solving task
vec <- c(temperature, meters_above_sea_level, auto, highest_building, population, area, shoppin_mall)

vec

vec_seq <- seq( from = 20, to = 100, by = 6)
vec_seq
  
vec_log <- sample(c(TRUE, FALSE), 10, replace = TRUE)
vec_log
```
## Question 3 Marix

Question description

Create a 3×3 matrix with random numbers between -50 and 50
(using sample() or runif()), and name its rows and columns according to your theme.

Then calculate row and column averages using rowMeans() and colMeans().
 


```{r}
 # code for solving task
m <- matrix(sample(-50:50, 9), nrow = 3, ncol = 3)
m

rowMeans(m)

colMeans(m)
```
## Question 4 List

Question description

Create a list that combines all your previous objects:
atomic variables, vectors, and matrix. Add one extra element — a short description of your universe as a string.


```{r}
# code for solving task
list <- list(auto, name, area, has_water, vec, m)
list
```
## Question 5 Factor

Question description

Create a factor variable representing categories within your universe, e.g.:

```{r}
# code for solving task

my_factor <- factor(c("big", "big", "small", "medium", "small"), levels = c("small", "medium", "big"))
my_factor
```
## Question 6 Data Frame

Question description

Create a data frame related to your object that includes:

5–8 observations (rows)
At least 3 columns of different data types
(e.g., name, numeric value, logical flag, category)
Then:

Rename the columns using names() or colnames().
Display its structure with str() and its summary with summary().
Add a new column derived from the others (for example, a ratio or condition).

```{r}
data <- data.frame(name=c("centr","podol", "almaznyj", "levada", "maka_scena"), numeric = c(16.4, 14.6, 25.7, 25.1, 18.2), logical = c(T, F, F, T, T), category = c("big", "small", "medium", "medium", "small"))

colnames(data) <- c("name", "size", "has_mall", "size")

data

str(data)

summary(data)

data$size_ratio <- data$size / max(data$size)

data

```

