---
title: "Task1"
author: "Roman Havrylenko"
date: "10/01/2024"
format:
  html:
    embed-resources: true
editor: source
---

## Question 1 Atomic Types

Create the following variables related to your chosen object:

4 numeric values (e.g., population, temperature, size, rating)
3 integer values (e.g., number of moons, buildings, inhabitants)
2 logical values (e.g., has_water, is_inhabited)
3 character values (e.g., name, region, category)

My object is Poltava city

```{r}
# code for solving task

#numeric
temperature <- 14.3
meters_above_sea_level <- 140.0
auto <- 285.000
highest_building <- 44.0


# integer

population <- 279593
area <- 103 #m2
shoppin_mall <- 3

#logical

has_water <-TRUE
has_autobus_station <- TRUE

#character

name <- 'Poltava'
region <- 'Poltavsky'
category <- 'city'

```


## Question 2 Vectors

Question description

Create three different vectors:

Numeric vector combining all your numeric and integer variables.
(Use c().)

Sequence vector from any number between 10–50 to 100, Choose step between 2 and 10.
(Use seq() or :.)

Logical vector of length 10 with random TRUE/FALSE values using sample().

```{r}
# code for solving task
vec <- c(temperature, meters_above_sea_level, auto, highest_building, population, area, shoppin_mall)

vec

vec_seq <- seq( from = 20, to = 100, by = 6)
vec_seq
  
vec_log <- sample(c(TRUE, FALSE), 10, replace = TRUE)
vec_log
```
## Question 3 Marix

Question description

Create a 3×3 matrix with random numbers between -50 and 50
(using sample() or runif()), and name its rows and columns according to your theme.

Then calculate row and column averages using rowMeans() and colMeans().
 


```{r}
 # code for solving task
m <- matrix(sample(-50:50, 9), nrow = 3, ncol = 3)
m

rowMeans(m)

colMeans(m)
```
## Question 4 List

Question description

Create a list that combines all your previous objects:
atomic variables, vectors, and matrix. Add one extra element — a short description of your universe as a string.


```{r}
# code for solving task
list <- list(auto, name, area, has_water, vec, m)
list
```
## Question 5 Factor

Question description

Create a factor variable representing categories within your universe, e.g.:

```{r}
# code for solving task

my_factor <- factor(c("big", "big", "small", "medium", "small"), levels = c("small", "medium", "big"))
my_factor
```
## Question 6 Data Frame

Question description

Create a data frame related to your object that includes:

5–8 observations (rows)
At least 3 columns of different data types
(e.g., name, numeric value, logical flag, category)
Then:

Rename the columns using names() or colnames().
Display its structure with str() and its summary with summary().
Add a new column derived from the others (for example, a ratio or condition).

```{r}
data <- data.frame(name=c("centr","podol", "almaznyj", "levada", "mala_scena"), numeric = c(16.4, 14.6, 25.7, 25.1, 18.2), logical = c(T, F, F, T, T), category = c("big", "small", "medium", "medium", "small"))

colnames(data) <- c("name", "size", "has_mall", "size")

data

str(data)

summary(data)

data$size_ratio <- data$size / max(data$size)

data

```
Task 2
2.1. Querying Vectors
Let’s start with the vectors you created in Task 1.

2.1.1. From your Sequence vector (the one with all your numbers):

Select and print the 3rd element.
Select and print elements 2 through 5.
Select and print all values that are greater than 20 (or another number that makes sense for your data).
2.1.2. From your Logical vector:

Select and print only the TRUE values. (Hint: your_vector[your_vector]).
Use which() to get the indices (positions) of all the TRUE values.
```{r}

vec_seq[3]
vec_seq[2:5]
vec_seq[vec_seq > 20]

vec_log[vec_log]
which(vec_log)
```
2.3. Investigating Your Data Frame
Use the data frame you created in Task 1.

Print the first 3 rows. (Use head().)
Print the value from the 4th row, 1st column.
Print only the column that contains your logical values (e.g., is_habitable) as a vector.
Logical Subsetting: Create a new data frame df_subset that contains only the rows that meet a specific logical condition.

Example: df_subset <- your_df[your_df$population > 1000, ]
%in% Subsetting: Create a new data frame df_subset_2 that selects rows belonging to two specific categories from your factor or character column.

Example: df_subset_2 <- your_df[your_df$region %in% c("North", "South"), ]
```{r}

head(data, 3)


data [4, 1]

data$has_mall


df_subset <- data[data$size > 20, ]
df_subset


df_subset_2 <- data[data$category %in% c("small", "medium"), ]
df_subset_2

```
2.4. Handling “Corrupted” Data (NA)
Create a copy of your original data frame (e.g., df_corrupted <- your_df).

Introduce 3 NAs into your df_corrupted at specific locations.

Example: df_corrupted[2, 3] <- NA
Find NAs:

Use sum(is.na(df_corrupted)) to find the total number of missing values.
Use colSums(is.na(df_corrupted)) to see how many NAs are in each column.
Calculate with NAs:

Calculate the mean() of the numeric column where you added an NA.
Filter NAs:

Use na.omit(df_corrupted) to create a new, “clean” data frame that contains only the complete rows.
```{r}

df_corrupted <- data

df_corrupted[2, 2] <- NA
df_corrupted[5, 3] <- NA
df_corrupted[3, 4] <- NA

df_corrupted

sum(is.na(df_corrupted))
colSums(is.na(df_corrupted))

mean(df_corrupted$size)
mean(df_corrupted$size, na.rm = TRUE)

df_clean <- na.omit(df_corrupted)
df_clean

```
2.5. Bonus Challenge: which()
Use which.max() or which.min() to find the row number (index) of the observation with the highest or lowest value in one of your numeric columns.

Use that index to print the row with that observation.

Example: highest_pop_index <- which.max(your_df$population)
Example: your_df[highest_pop_index, ]
```{r}
highest_size_index <- which.max(data$size)
data[highest_size_index, ]

lowest_size_index <- which.min(data$size)
data[lowest_size_index, ]

```
